\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-06-19 paper/paper.tex
Consistent usage of "drive-model" to distinguish from many other uses of the word model

Failure data for series systems may be collected at the system level
by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
model (GLFP), which provides a generative model for
failure times in a population where some units are susceptible to infant mortality. This model
covers a subset of the class of distributions known as bathtub distributions
\cite{sujata}. Bathtub distributions are characterized by a U-shaped
hazard function, where early failures are due to infant mortality and
late failure are due to wearout. While complex systems actually
exhibit more than two modes of failure, suitable parametric models
which can suitably approximate the overall lifetime distribution are
useful.

@@

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GLFP lifetime model.  Product populations often contain a mixture of defective and non defective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{sujata}.  The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}. \\

While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.  A naive approach would combine all drive models and fit an overall GLFP model.  This model can be easily estimated, but a common GLFP model for all drive brand groups fits poorly: the model captures drive models with lots of observed failures, while the predicted time to failure for drive brand models with less information is over or under estimated.  At the other extreme, fitting individual models to each hard drive brand excludes many drive brand models with few failures.  The GLFP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.  \\

We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of the GLFP parameters for each drive model group.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive brand models to obtain stable estimates.  This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.

\subsection{Overview}
The structure of the paper is as follows.  Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.  Section 2 introduces the GLFP model as a mixture of two Weibull distributions.  In Section 3 we fit multiple specifications of the GLFP model to best describe the Backblaze data.  We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.  In Section 4 we compare the different model specifications, in addition to the hard drive brand models based on our final GLFP model.  Finally, in Section 5, we review the GLFP model and propose future extensions.

\section{Data}
Backblaze is a company that offers backup storage to protect against data loss.  While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive brand models.  The hard drives continuously spin in controlled storage pods where they run until failure.  When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.  In addition, the number of storage pods is increasing as Backblaze adds new hard drive models to the sample.  Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive models currently operating.  No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.

As of the first quarter of 2016, Backblaze was collecting data on 63 different hard drive models.  Some drive models have been running since 2013, while others were added at a later date.  There are a total of 75,297 hard drives in operation.  The distribution of drives by model varies: some models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.  Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive brand models with at least 3 failures.  For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.

@@

Applying this method to the hard drive data, we first must estimate
the empirical cdf for each model using the Kaplan-Meier estimator
\cite{kaplan}.  With left truncation, however, the standard
Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by
Turnbull, and given in more detail by Meeker and Escobar
\cite{turnbull,meeker}.  For each hard drive model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by

@@

In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model 14 on Weibull paper.  Drive model 14 had the most observed failures of all the brands.  Each point on the plot corresponds to an observed failure.  Censored drives are not plotted.  Standard error bands are calculated using Greenwood's method \cite{green}.  As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.  One mode is a result of manufacturing defects, which cause early failures, known as infant mortality.  The second mode is non-defective hard drives that eventually fail due to wearout.   Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring between 8000 and 20,000 hours.  Therefore, fitting a single Weibull model is not flexible enough to model the failure distribution.

\begin{figure}[H]
\centering
\includegraphics[width=.5\textwidth]{fig/mod14-KM-adj}
\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
\end{figure}

\section{Hierarchical GLFP model}
For modeling the lifetime of hard-drives, we select the Generalized Limited Failure Population model of \citet{chan}.
Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$

@@

To borrow strength across models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters. For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:

@@

The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations. Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.  Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by brand drive model.

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive brand models. We found that there was insufficient information in the data to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).

@@

\begin{enumerate}
\item This model constrains all drive models to the same failure distribution. For this ``reduced" model we assume the priors :

@@

The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit. The upper left panel shows the estimated failure curve fit to all data ignoring drive model as a factor (Model 1). The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.

The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which mean that the ranking of models with respect to fraction failing differs as a function of time.

Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.

@@

We can also compare the model fits for each drive brand model individually.  Using the set of parameters that maximize the posterior log probability, we examine the GLFP cdf for each of the 4 model specifications.  The GLFP fit for drive brand models 2, 9, 14, and 40 are presented in Figure 4.  The black step function again corresponds to the adjusted Kaplan-Meier estimates.  As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.  All the GLFP models overestimate the proportion of failed hard drives for drive model 2, but Model 4 is the least biased.  For drive model 14, Models 1-3 underestimate the proportion of drives failing, and for drive model 40, they slightly overestimate the data.  Conversely, Model 4 follows the observed failures quite closely--accurately capturing failures at the lower and upper ends of the distribution.  The difference in accuracy between Models 3 and 4 are and Models 1 and 2 are visually quite apparent.  Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive model 40.
\begin{figure}[H]
    \centering
   \includegraphics[width=\textwidth]{fig/mod_compare_legend}
		\caption{Estimated GLFP for 4 Drive Brand Models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
\end{figure}

@@

We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.  On the left panel of Figure 5 we plot B10, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive brand models. The best six drive models ranked by TTF are also the drive models with the longest time to B10.  Drive brand models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.  Another interesting feature to compare is $\pi$, the proportion failing due to infant mortality.  One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each hard drive model.  The plot is again sorted according to B10 so it can be compared to the left hand side.  While for many of the drive models the ordinal ranking is the same as on the left, there are some drive model comparisons, for example 23 and 18, that differ in ranking if compared using B10 or $\pi$.  Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.

@@

\section{Conclusions}
We should point out some assumptions made in this analysis. First, we assume exchangability of units within drive models. This assumption is due to our ignorance about the causes of failure and of potentially important covariates. If we had information about distinguishing characteristics of the two identified failure modes, then we should use this information. Also, there could be many environmental or human-related factors that impact lifetime, such as operating conditions associated with location in the data center, which are likely confounded with drive-model.

Second, it is probably true that many modes of failure occur and that the heterogeneity that we observe among drive models is due to different combinations of these. Our ignorance of the causes of failure leads us to approximate the main failure mode with a single drive-model specific Weibull distribution. Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our ``model certainty" makes the uncertainty about these parameters too small. On the other hand, our model borrows strength across drive-models, including those with information about late failures. In addition, we suggest that using metrics for comparing drive-models that reduce the importance of the right tail of the distribution are appropriate for these data. With respect to decisions based on such metrics, we believe the assumptions in the previous paragraph to be the most critical. Those issues are typical to observational data such as these.

@@

Analysis using HMC is accessible to practitioners by Stan. HMC can be very efficient in terms of effective samples because of the way that it uses joint updates for all parameters using Hamiltonian dynamics to rapidly explore the target distribution.

>>>>

Failure data for series systems may be collected at the system level
by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
model (GLFP), for
failure times in a population where some units are susceptible to infant mortality. This model
covers a subset of the class of distributions known as bathtub distributions
\cite{sujata}. Bathtub distributions are characterized by a U-shaped
hazard function, where early failures are due to infant mortality and
late failure are due to wearout. While complex systems actually
exhibit more than two modes of failure, suitable parametric models
which can suitably approximate the overall lifetime distribution are
useful.

@@

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GLFP lifetime model.  Product populations often contain a mixture of defective and non defective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive-models \cite{sujata}.  The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}. \\

While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.  A naive approach would combine all drive-models and fit an overall GLFP model.  This model can be easily estimated, but a common GLFP model for all drive-models fits poorly: the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.  At the other extreme, fitting individual models to each drive-mode is difficult for drive-models with few failures.  Due to multimodality, in these cases the model fit is unstable unless strongly informative priors are chosen.  \\

We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.  This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.

\subsection{Overview}
The structure of the paper is as follows.  Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.  Section 2 introduces the GLFP model as a mixture of two Weibull distributions.  In Section 3 we fit multiple specifications of the GLFP model to best describe the Backblaze data.  We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.  In Section 4 we compare the different model specifications, in addition to the hard drive-models based on our final GLFP model.  Finally, in Section 5, we review the GLFP model and propose future extensions.

\section{Data}
Backblaze is a company that offers backup storage to protect against data loss.  While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods where they run until failure.  When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.  In addition, the number of storage pods is increasing as Backblaze adds new drive-models to the sample.  Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.

As of the first quarter of 2016, Backblaze was collecting data on 63 different drive-models.  Some drive-models have been running since 2013, while others were added at a later date.  There are a total of 75,297 hard drives in operation.  The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.  Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive-models with at least 3 failures.  For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.

@@

Applying this method to the hard drive data, we first must estimate
the empirical cdf for each drive-model using the Kaplan-Meier estimator
\cite{kaplan}.  With left truncation, however, the standard
Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by
Turnbull, and given in more detail by Meeker and Escobar
\cite{turnbull,meeker}.  For each drive-model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by

@@

In Figure 2 we plot the Kaplan-Meier adjusted cdf for drive-model 14 on Weibull paper.  Drive-model 14 had the most observed failures of all the brands.  Each point on the plot corresponds to an observed failure.  Censored drives are not plotted.  Standard error bands are calculated using Greenwood's method \cite{green}.  As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.  One mode is a result of manufacturing defects, which cause early failures, known as infant mortality.  The second mode is non-defective hard drives that eventually fail due to wearout.   Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring between 8000 and 20,000 hours.  Therefore, fitting a single Weibull model is not flexible enough to model the failure distribution.

\begin{figure}[H]
\centering
\includegraphics[width=.5\textwidth]{fig/mod14-KM-adj}
\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive-model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
\end{figure}

\section{Hierarchical GLFP model}
For modeling the lifetime of hard drives, we select the Generalized Limited Failure Population model of \citet{chan}.
Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$

@@

To borrow strength across drive-models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters. For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:

@@

The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations. Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.  Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by drive-model.

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models. We found that there was insufficient information in the data to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).

@@

\begin{enumerate}
\item This model constrains all drive-models to the same failure distribution. For this ``reduced" model we assume the priors :

@@

The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit. The upper left panel shows the estimated failure curve fit to all data ignoring drive-model as a factor (Model 1). The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive-models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.

The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which mean that the ranking of drive-models with respect to fraction failing differs over time.

Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive-model, there is greater variability in the early rate of failure and, for some drive-models, the forecasted fraction failing is adjusted downward.

@@

We can also compare the model fits for each drive-model individually.  Using the set of parameters that maximize the posterior log probability, we examine the GLFP cdf for each of the 4 model specifications.  The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure 4.  The black step function again corresponds to the adjusted Kaplan-Meier estimates.  As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.  All the GLFP models appear to overestimate the proportion of failed hard drives for drive-model 2, but Model 4 agrees the most with the nonparametric estimate.  For drive-model 14, Models 1-3 underestimate the proportion of drives failing, and for drive-model 40, they slightly overestimate the data.  Conversely, Model 4 follows the observed failures quite closely--accurately capturing failures at the lower and upper ends of the distribution.  The difference in accuracy between Models 3 and 4 are and Models 1 and 2 are visually quite apparent.  Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive-model 40.
\begin{figure}[H]
    \centering
   \includegraphics[width=\textwidth]{fig/mod_compare_legend}
		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
\end{figure}

@@

We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.  On the left panel of Figure 5 we plot B10, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models. The best six drive-models ranked by TTF are also the drive-models with the longest time to B10.  Drive-models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.  Another interesting feature to compare is $\pi$, the proportion failing due to infant mortality.  One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each drive-model.  The plot is again sorted according to B10 so it can be compared to the left hand side.  While for many of the drive-models the ordinal ranking is the same as on the left, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using B10 or $\pi$.  Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.

@@

\section{Conclusions}
We should point out some assumptions made in this analysis. First, we assume exchangability of units within drive-models. This assumption is due to our ignorance about the causes of failure and of potentially important covariates. If we had information about distinguishing characteristics of the two identified failure modes, then we should use this information. Also, there could be many environmental or human-related factors that impact lifetime, such as operating conditions associated with location in the data center, which are likely confounded with drive-model.

Second, it is probably true that many modes of failure occur and that the heterogeneity that we observe among drive models is due to different combinations of these. Our ignorance of the causes of failure leads us to approximate the main failure mode with a single drive-model specific Weibull distribution. Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our uncertainty in the right tail of the distribution is probably too small. On the other hand, our model borrows strength across drive-models, including those with information about late failures. In addition, we suggest that using metrics for comparing drive-models that reduce the importance of the right tail of the distribution are appropriate for these data. With respect to decisions based on such metrics, we believe the assumptions in the previous paragraph to be the most critical. Those issues are typical to observational data such as these.

@@

Analysis using HMC is accessible to practitioners by Stan. HMC has been shown to be very efficient in effective samples per iteration because its joint updating scheme.
